---
title: "Analysis"
---
Before we dive into it, let us explain what each variable means:
Player: Refers to either Messi or Ronaldo.
Season: Indicates the season when the goal was scored. This matters because each season is different in many factors; neither Ronaldo nor Messi remains the same player throughout their football career.
Competition: Indicates the competition in which the goal was scored. Some competitions are more significant than others, and some leagues are more competitive or harder to perform well than others.
Matchday: Indicates which matchday the goal was scored in. A later matchday or round indicates a later time in the competition, which bears significance as well. A later match in a round has a higher weightage than an earlier one. 
Date: Indicates the date of the goal scored.
Venue: Indicates whether the goal was scored in a Home (H) or an Away (A) game. Generally, A games are harder to perform in because it is not your home stadium and a majority of the fans there would be cheering for the other team – adding to the mental challenge. 
Club: Refers to the club that the goal was scored for.
Opponent:  Refers to the opponent club that the goal was scored against.
Result: Refers to the final score of the match.
Playing Position: Refers to which position on the field the player was playing. Generally, it is easier to score from a central position. 
Minute: Refers to the minute of the match at which the goal was scored.
Score: Refers to the score tally when the goal was made.
Type: Refers to the type of goal scored. Different types of goals correspond to differing difficulties of execution.
Goal assist: Refers to the player who assisted the goal. Some players are better skilled than others, hence receiving a ball from these players increases the possibility of one scoring. 


## General Observations

Before we analyse the plots, it is always good practice to see what observations we can make from the dataset itself. 

Below are some summary statistics about the dataset. 

```{r, echo=F, message=F, warning=F}
library("tidyverse")
library(dplyr)
```


```{r, eval=TRUE,echo=FALSE, warning=F, message=F}
goatdebate <- read.csv("goat.csv")

total_goals_by_player <- 
  goatdebate %>%
  group_by(Player) %>%
  summarise(TotalGoals = n())

print(total_goals_by_player)

average_goals_per_season <- goatdebate %>%
  group_by(Player, Season) %>%
  summarise(TotalGoals = n()) %>%
  group_by(Player) %>%
  summarise(AverageGoalsPerSeason = mean(TotalGoals))

print(average_goals_per_season)

```
As one can observe from this summary, Ronaldo’s total goals tally up to 2 more than Messi’s. However, Messi has played for a lesser number of seasons, consequently meaning that he had a shorter career than Ronaldo. Also, on average, Messi has scored more goals per season than Ronaldo. Thus, on the surface and based on these statistics, it would seem conclusive that Messi is a better player. However, let us examine their respective performances more deeply within certain categories!

## Interactive Shiny App

```{=html}
<iframe height="400" width="100%" frameborder="no" src="https://amirnotsosmart.shinyapps.io/sample_app_nm2207/"> </iframe>
```


## Goals Over Seasons - Time Plot

Firstly, let us analyse the time plot indicating the goals over seasons. Ronaldo has played for 25 seasons so far while Messi has played for 19. One could argue that Messi is better because his highest-scoring season is above 70 goals. Meanwhile, Ronaldo did not break that threshold, with his highest-scoring season only having around 60 goals.

However, Ronaldo had 2 seasons crossing the threshold of 60 or more goals, while Messi only had one season – his highlight season of above 70 goals. Ronaldo also had 1 more season where he scored more than 50 goals, with 6 while Messi had only 5, and Ronaldo had only 5 seasons in 25 seasons where he scored less than 20 goals, while Messi had 6 seasons scoring less than 20 goals. Hence, statistically, although Messi experienced a higher performance maximum, Ronaldo can be argued to be a much more consistent, and well-performing scorer across his football career.

## Goals by Type 

Second, from the goals by type plot, Ronaldo has scored in a larger variety of ways than Messi. Ronaldo is renowned for his dominant right foot, while Messi is well-known for being a left-footed player which is rather rare in football. Between both players, Ronaldo has scored more goals with his non-dominant left foot than Messi has with his right. Ronaldo has also scored more goals with his head than Messi. Thus, Ronaldo’s versatile proficiency as opposed to Messi adds layers of unpredictability to his goal-scoring repertoire.

## Goals By Position

Next, we analyse the goals by position. Ronaldo has scored slightly over 200 goals from the Centre Forward (CF) position, over 350 goals from the Left Wing (LW), and nearly 100 goals from the Right Wing (RW). Meanwhile, Messi has scored over 250 goals from CF, over 200 goals from RW, and an almost negligible number of goals from LW. 

Although Messi’s negligibility of goals from LW may be due to playing fewer games in that position, the statistics continue to be in favour of Ronaldo as a more versatile scorer. Ronaldo has scored more goals from the sides as compared to Messi, whose goals are mainly concentrated through the middle. In football, scoring successfully from the wings requires a greater amount of tactility and skill as opposed to the middle. Despite acknowledging possible reasons behind Messi’s negligibility of goals from LW, in sum, Ronaldo still scores a greater total of goals from wing positions as opposed to Messi – another testament to his athletic capabilities. 


## Goals by Minute Heatmap

Further, let us analyse the goal distribution by the match minute through a heatmap. Ronaldo has scored more goals than Messi after the 90th minute, the most crucial moment of any football match, thus pointing to him as a clutch scorer. As the minutes in the match drag on, players get more fatigued making it harder to score.

While the fatigue builds up, so does the pressure of the game. Hence, it takes a great player who can excel under pressure to accurately score in the later stages of the game. However, aside from goals scored after the 90-minute mark, admittedly both Messi and Ronaldo have a very high percentage of goals late into the game within 90 minutes.

Thus, both can be argued to be clutch players. Regardless, this metric is tough to compare singularly between because, within 90 minutes, the goals scored bear differing significances. For example, a goal scored in a 4-1 match difference weighs significantly less than one scored as the winning goal in a 4-4 match. 

## Some additional plots for analysis 

```{r, eval=TRUE,echo=FALSE}
library("tidyverse")
library(dplyr)
goatdebate <- read.csv("goat.csv")

goals_by_league <- goatdebate %>%
  group_by(Player, Competition) %>%
  summarise(Goals = n(), .groups = 'drop')

goals_by_league %>%
  pivot_wider(names_from = Player, values_from = Goals) %>%
  na.omit() %>%
  pivot_longer(`Cristiano Ronaldo`:`Lionel Messi`, names_to = "Player", values_to = "Goals") %>%
  mutate(Competition = reorder(Competition, -Goals)) %>%
  ggplot( aes(x = Competition, y = Goals, fill = Player)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Number of Goals in Each League by Player",
       x = "League",
       y = "Number of Goals") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_log10()

goals_by_league %>%
  filter(Player == "Cristiano Ronaldo") %>%
  mutate(Competition = reorder(Competition, -Goals)) %>%
  ggplot( aes(x = Competition, y = Goals)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Number of Goals in Each League for Ronaldo",
       x = "League",
       y = "Number of Goals") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_log10()

goals_by_league %>%
  filter(Player == "Lionel Messi") %>%
  mutate(Competition = reorder(Competition, -Goals)) %>%
  ggplot( aes(x = Competition, y = Goals)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Number of Goals in Each League for Messi",
       x = "League",
       y = "Number of Goals") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_log10()
```

Above is a plot that indicates Messi and Ronaldo’s goals in the respective leagues they have played in. Different leagues in football are accompanied by different playing styles, settings, and competitiveness. Ronaldo has played and scored in more leagues than Messi, indicating that he is a more versatile player because he can adapt and perform well despite different settings. 

Furthermore, he has played in the Premier League – the highest level of the English football league system and a highly competitive, global league – while Messi has not. Furthermore, he has more goals in the Champions League,  deemed the best competition in the world. Interestingly, Ronaldo has scored a total of 60 goals across 49 Champions League games – a record that nobody else has managed to come close to in football.

To make this more interesting, we can zoom in to each players' goals by their respective competitions through this app!

```{=html}
<iframe height="400" width="100%" frameborder="no" src="https://amirnotsosmart.shinyapps.io/app2r/"> </iframe>
```

```{r, eval=TRUE,echo=FALSE}
library("tidyverse")
library(dplyr)
goatdebate <- read.csv("goat.csv")

goatdebate$Result <- as.character(goatdebate$Result)

goat_data_wins <- goatdebate %>%
  separate(Result, into = c("GoalsFor", "GoalsAgainst"), sep = ":", convert = TRUE) %>%
  mutate(Win = if_else(GoalsFor > GoalsAgainst, 1, 0)) %>%
  group_by(Player) %>%
  summarise(Wins = sum(Win, na.rm = TRUE))

ggplot(goat_data_wins, aes(x = Player, y = Wins, fill = Player)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Wins by Player", x = "Player", y = "Wins") +
  theme_minimal()
```

Above is a plot showing the number of wins for each player. While Ronaldo has slightly more wins, it is important to note that Messi has played fewer seasons than him. Consequently, Messi likely has a higher average win percentage than Ronaldo, despite Ronaldo having more wins. 

## Summary

1. Total Goals Tally: While Ronaldo has scored slightly more goals overall, Messi's higher goals-per-season average suggests exceptional consistency and efficiency in front of the goal.

2. Goals by Type: Ronaldo's versatility in scoring, including goals with his weaker foot and head, showcases a broader skill set compared to Messi, adding unpredictability to his gameplay.

3. Goals by Position: Ronaldo's ability to score from various positions, especially from the wings, demonstrates his adaptability and effectiveness across different areas of the field, contrasting with Messi's concentration of goals through the middle.

4. Goal Distribution by Match Minute: Ronaldo's tendency to score crucial goals, especially after the 90th minute, suggests his capability as a clutch player, although both players exhibit a high percentage of late-game goals.

5. League Performance: Ronaldo's success in multiple leagues, including the highly competitive Premier League, highlights his adaptability and ability to excel in different football environments, further giving him an edge in terms of versatility.



